Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids

### 2019-10-09 Manuscript/manuscriptPGPE.tex
Update text and figures

\subsection{Counterion binding to PE and PG lipid bilayers}
In our previous study about PS lipids \cite{antila18}, the underestimated increase of PC
headgroup order parameters upon addition of negaticely charged lipids
was related
to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
Similar correlation is observed here in %other than GROMOS-CKP simulation:
Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
and larger decrease in POPC $\beta$ carbon order parameter upon addition of POPG than other simulations.
Together with the lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$),
the results suggest that the counterions bind too strongly and shield the electrostatic repulsion between PG headgroups
in bilayers simulated with Lipid17 parameters.
In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of POPC
increases upon addition of POPG more than in CHARMM36 and Slipids simulations even though the binding affinities in these
simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
and other decreases \todo{Can we figure out the reason for this?}.
In conclusion, the counter-ion binding seems to be overestimated in Lipid17 with the strongest affinity among the
tested models, while with this data we cannot fully evaluate binding affinity in the other models.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
  \caption{\label{CIdensPG}
    Counterion densities and area per lipids from simulations with PG lipids.
    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
  }
  \todo{Data for GROMOS is also available without ions, but I could not analyze it, probably because tpr file was not correct.}
\end{figure}

%Sodium binding to PC lipid bilayers is significantly overestimated by most simulations models \cite{catte16}.
%Also, sodium binding to PS lipids seems to be overestimated, although the
%presence of counterions complicate the comparison for negatively charged lipids \cite{NMRlipidsIVps}.
Similarly to POPG, the Lipid17 exhibits stronger sodium binding affinity
to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
Experimental data for sodium binding to PE lipids is not available, but
large deviations from PC would be surprising.
However, in Slipids and GROMOS-CKP simulations, the binding affinity to POPE is much weaker than to POPC
(our previous work \cite{catte16} and Fig. ??). On the other hand, 
the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
previous Amber lipid version, Lipid14.
\todo{The potential differences in the used ion models should be checked}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
  \caption{\label{NAdensPE}
    Sodium (solid line) and choride ion density profiles along membrane normal
    from different simulations with PE lipids.
  }
\end{figure}

>>>>

\subsection{Sodium binding to PE and PG lipid bilayers}
In our previous study about PS lipids \cite{antila18}, the underestimated increase of PC
headgroup order parameters upon addition of negaticely charged lipids
was related
to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
Similar correlation is observed here in %other than GROMOS-CKP simulation:
Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
and larger decrease in POPC $\beta$ carbon order parameter upon addition of POPG than other simulations.
Together with the lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$),
the results suggest that the counterions bind too strongly and shield the electrostatic repulsion between PG headgroups
in bilayers simulated with Lipid17 parameters.
In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of POPC
increases upon addition of POPG more than in CHARMM36 and Slipids simulations even though the binding affinities in these
simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
and other decreases \todo{Can we figure out the reason for this?}.
In conclusion, the counter-ion binding seems to be overestimated in Lipid17 with the strongest affinity among the
tested models, while with this data we cannot fully evaluate binding affinity in the other models.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
  \caption{\label{CIdensPG}
    Counterion densities and area per lipids from simulations with PG lipids.
    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
  }
  \todo{Data for GROMOS is also available without ions, but I could not analyze it, probably because tpr file was not correct.}
\end{figure}

%Sodium binding to PC lipid bilayers is significantly overestimated by most simulations models \cite{catte16}.
%Also, sodium binding to PS lipids seems to be overestimated, although the
%presence of counterions complicate the comparison for negatively charged lipids \cite{NMRlipidsIVps}.
Similarly to POPG, the Lipid17 exhibits stronger sodium binding affinity
to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
In Lipid17 the sodium binding affinity to POPE is approximately similar as to POPC,
while in other models exhibit weaker binding to POPE (Fig. \ref{NAdensPC}).
The difference is small in Slipids and CHARMM36, while GROMOS-CKP predicts
subtantially stronger binding to POPC.
Experimental data for sodium binding to PE lipids is not available, while
simulations in agreement with electrometer data from NMR suggests that
sodium binding to POPC is weaker than in any of the simulations
here (Fig. \ref{NAdensPC})~\cite{catte16,melcr18}. Assuming that the binding
to POPE would be similar than to POPC, the sodium binding affinity to POPE
is potentially realistic in CHARMM36, Slipids, and GROMOS-CKP simulations here,
but substantially overestimated in Lipid17 simulation.
%On the other hand, the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
%previous Amber lipid version, Lipid14.
\todo{The potential differences in the used ion models should be checked}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
  \caption{\label{NAdensPE}
    Sodium (solid line) and choride ion density profiles along membrane normal
    from different simulations with PE lipids.
  }
\end{figure}

### 2019-10-09 Manuscript/manuscriptPGPEsuppl.tex
Update text and figures

\clearpage
\bibliography{refs.bib}

>>>>

\clearpage

\section{Changes of PG headroup order parameters upon addition of PC}

\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPC.eps}
  \caption{\label{HGorderparametersPGvsPC}
    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
  }
\end{figure}

\clearpage

\section{Sodium binding to POPC simulations}

\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/NAdensPC.eps}
  \caption{\label{NAdensPC}
    Sodium (solid line) and choride ion density profiles along membrane normal
    from different simulations with PC lipids.
  }
  \todo{Discussion about differences to the NMRlipids II to be discussed once we have the details on ions models.}
\end{figure}

\bibliography{refs.bib}
